Cycles of the logistic map

### 2013-05-15 doc/pc-ijbc.tex
add a few references, hengb.magma scripts
sightly improve doc

We will illustrate the algorithm on the logistic map \cite{may, strogatz}
  defined in \refeq{logmap}.
If $r$ is real,
  we will find windows $(r_a, r_b)$,
    within which stable $n$-cycles can exist.
%
%The window boundaries can be classified
%  to onset and bifurcation points according to \refeq{der};
Here, if $r > 0$, then
  $r_a$ ($r_b$) are the onset (bifurcation) points.
%
There are generally multiple such windows for a single $n$,
  but all onset points satisfy the same polynomial equation,
  and all bifurcation points another.
%
Our goal is thus to find the two polynomials for a given $n$.

@@

First, to count the square-free cyclic polynomials,
  we notice a one-to-one mapping
  from the square-free cyclic polynomials
  to the binary necklaces.
%
A binary necklace
  is a nonequivalent binary $0$-$1$ string.
%
Two strings are equivalent if they differ only by a circular shift.
%
For example, for $n = 3$, there are $2^3 = 8$ binary strings,
but only four necklaces: 000, 001, 011 and 111,
since 010 and 100 are equivalent to 001,
so are 110 and 101 to 011.
%
The mapping is constructed as follows.
%
Given a cyclic polynomial, if the generator contains $x_k$,
  the $k$th character from the right of the necklace is 1,
  otherwise 0,
e.g. the above four necklaces correspond to
  $C_0$, $C_1$, $C_{12}$ and $C_{123}$, respectively.
%For example, if $n = 3$,
% the square-free polynomials for $000$, $100$, $110$ and $111$ are
%  $C_0 = 1$ (generator: 1),
%  $C_1 = x_1 + x_2 + x_3$ (generator: $x_1$),
%  $C_{12} = x_1 x_2 + x_2 x_3 + x_3 x_1$ (generator: $x_1 x_2$)
%  and $C_{123} = x_1 x_2 x_3$ (generator: $x_1 x_2 x_3$), respectively.
Thus,
%
%\begin{theorem}
the number $\NB(n)$ of the square-free cyclic polynomials $C_p$,
  %$(\vx)$ formed by the cycle points $\vx = \{x_1, \ldots, x_n\}$,
  is given by
\begin{equation}
  %\deg_\lambda A_n\lam =
  \NB(n) = (1/n) \textstyle\sum_{d|n} \varphi(n/d) 2^{\,d},
\label{eq:necklace}
\end{equation}
%
%\label{thm:necklace}
%\end{theorem}
%
where $\varphi(m)$ is Euler's totient function,
which gives the number of integers from 1 to $m$
  that are coprime to $m$,
  e.g.,
  $\varphi(1) = 1$,
  $\varphi(3) = 2$ for 1 and 2,
  $\varphi(6) = 2$ for 1 and 5.
By \refeq{xr},
  $\NB$ is also $\deg_\lambda A_n$.
The first few $\NB(n)$ are
2, 3, 4, 6, 8, 14, 20, 36, 60, 108, \dots, % 188, 352, 632, \ldots,
  from $n = 1$.

@@

Incidentally, $\LB$ is also the number of aperiodic necklaces of length $n$
(e.g., for $n = 4$,
  ``$0011$'' is aperiodic, but ``$0101$'' is not due to the shorter period 2).
%
Since an $n$-necklace is either aperiodic itself or
  a repeated aperiodic $d$-necklace (with $d|n$),
  we have
\begin{equation}
  \NB(n) = \textstyle\sum_{d|n} \LB(d).
\label{eq:necklacelyndon}
\end{equation}
%
Since the left side is the degree in $\lambda$ of $A_n\lam$,
  which accommodates all fixed points of $f^n$,
  %\refeq{necklacelyndon} shows that
  each distinct $d$-cycle (with $d|n$)
  contributes one and only one zero to $A_n\lam$.
%
%
We may thus write
%\begin{theorem}
  \begin{equation}
    A_n\lam = \textstyle\prod_{d|n} Q_d^{n/d}\lam,
    \label{eq:prod}
  \end{equation}
where $Q_d^{n/d}(\lambda) \equiv \prod_l
  \big[\lambda - \Lambda\big(\vxn{n}^{(l)}\big) \big]$
  with $\vxn{n}^{(l)}$ being the $n$ points
  corresponding to the $l$th $d$-cycle.
%
In order words, $\vxn{n}^{(l)}$ is $\vxn{d}^{(l)}$
  repeated for $c = n/d$ times:
\begin{equation}
  \vxn{n}^{(l)} =
  \{
    x_1^{(l)}, \ldots, x_d^{(l)},
    \; \ldots, \;
    x_1^{(l)}, \ldots, x_d^{(l)}
  \}
  = \vxn{d}^{(l)} * c.
\notag
\end{equation}
%
It follows that
$\Lambda(\vxn{n})$ = $[\Lambda(\vxn{d})]^c$.
By factorizing $\lambda - \Lambda^c$
  in terms of complex roots of unity, we get
\begin{align}
Q_d^{c}\lam
%  &= \prod_l
%    \Big\{
%      \lambda - \big[\Lambda\big(\vxn{d}^{(l)}\big) \big]^{c}
%    \Big\}
%    \notag \\
  &=
    \prod_l
    \Big\{
      \prod_{k=1}^c \,
      \big[
          e^{2k\pi i/c} \lambda^{1/c}
          - \Lambda\big(\vxn{d}^{(l)}\big)
      \big]
    \Big\},
    \notag
\end{align}
%
Now changing the order of the two products yields
\begin{equation}
Q_d^{c}\lam
  =  \prod_{k=1}^c P_d\big(e^{2k\pi i/c} \lambda^{1/c}\big),
\label{eq:QP}
\end{equation}
%
%
Despite the presence of $\lambda^{1/c}$,
  $Q_d^c\lam$ is a polynomial of $\lambda$,
  thanks to the invariance
  $\lambda \rightarrow e^{2\pi i} \lambda$;
and
  $\deg_\lambda Q_d^c\lam
    = \deg_\lambda P_d\lam = \LB(d)$.

@@

We can therefore distinguish \emph{original} cycles
  from those born out of the above branching.
Similar to \refeq{QP},
  the contributions from the non-original cycles
  born out of the $d$-cycles
  can be written as a cyclotomic product
  $\prod_{(k, c) = 1}
      P_{d}
        \left(
          R, e^{2k\pi i/c}
        \right)$,
where $k$ runs through integers from 1 to $c=n/d$,
  that are coprime to $c$.
With these contributions removed,
  we obtain the polynomial $S_n(R)$ responsible for the original cycles:
  %or the \emph{original factor} below,
  \begin{equation}
    S_n(R)
    = \frac
    {
      P_n(R, 1)
    }
    {
      \prod_{c d =  n, \; c > 1}
      %\left[ \,
        \prod_{(k, c) = 1}
      P_{d}
        \left(
          R, e^{2k\pi i/c}
        \right)
      %\right]
    },
  \label{eq:origfac}
  \end{equation}
%
%
%
Take the $n = 4$ case as an example, we have
$\prod_{k=1,3} P_1(R, e^{k\pi i/2})
%=(-4R-2i-1)(-4R+2i-1)
=16R^2+8R+5$
(for $c = 4$),
$P_2(R, -1) = 4R - 5$
(for $c = 2$),
and
$P_4(R, 1)
= (4R - 5) (16 R^2 + 8 R + 5)
  \bigl[
    (4R - 3)^3 - 108
  \bigr]$.
Dividing $P_4(R, 1)$ by the two factors yields
$S_4(R) = (4R-3)^3 - 108$.
%  whose only real root $R_a=(3+\sqrt[3]{108})/4$ corresponds to
%  the onset of the original cycle.
Note $R_a' = 5/4$ is not original,
  for it is born out of period-doubling the 2-cycle.
The two cycles are compared in \reffigs{cobweb}(d) and (e).

@@

A simple two-dimensional extension of the logistic map
  is the H\'enon map \cite{henon}:
\begin{equation}
  x_{k+1} = 1 + y_k - a \, {x_k}^2, \quad
  y_{k+1} = b \, x_k.
\label{eq:henon}
\end{equation}
%
By a change of variables
  $x_k^\mathrm{new} \equiv a \, x_k^\mathrm{old},
   y_k^\mathrm{new} \equiv a \, y_k^\mathrm{old}$,
we have, in terms of the new variables,
\begin{equation}
  x_{k+1} = a + y_k - {x_k}^2, \quad
  y_{k+1} = b \, x_k.
%\tag{$\ref{eq:henon}'$}
\label{eq:henons}
\end{equation}
%
%Since neither $a$ nor $b$ is changed during the transformation,
\refeq{henon} and \refeq{henons} share the
same onset and bifurcation points
in terms of $a$ and $b$.
%
With $b \rightarrow 0$ and $a \rightarrow R$,
we recover the logistic map \refeq{logmaps}.

@@

The counterpart of \refeq{der}
  can be found from the Jacobian of $n$th iterate of \refeq{henons}
\[
  \vct J^n
  \equiv
  \left(
    \begin{array}{ccc}
      \partial x_{n+1}/\partial x_1 & \partial x_{n+1}/\partial y_1 \\
      \partial y_{n+1}/\partial x_1 & \partial y_{n+1}/\partial y_1 \\
    \end{array}
  \right),
\]
and the magnitude of both eigenvalues of the matrix must be less than 1.
By the chain rule, we have
  $\vct J^n (x_1) = \vct J^1(x_1) \cdots \vct J^1(x_n)$,
  where
$\vct J^1(x_k)
  =
  \left(
    \begin{array}{cc}
      -2 x_k & 1 \\
      b & 0
    \end{array}
  \right)$.
Thus, the eigenvalues $\lambda$ of $\vct J^n$ can be found from
%
\begin{equation}
\big|\, \lambda \, \vct I - \vct J^n(x_1) \,\big|
  = \lambda^2 - \Theta(\vx) \,\lambda +(-b)^n = 0,
%\tag{$\ref{eq:der}'$}
\label{eq:henonder}
\end{equation}
%
where $\Theta(\vx)$ and $(-b)^n$
  are the trace and determinant of $\vct J^n(x_1)$,
  respectively \cite{hitzl}.
The onset and bifurcation points
  correspond to $\lambda = +1$ and $-1$, respectively

@@

As an example of a higher-order %one-dimensional
map, we consider the following cubic map \cite{strogatz}
%
\begin{equation}
  x_{k + 1} = f(x_k) = r \, x_k - {x_k}^3.
\label{eq:cubic}
\end{equation}
%
Unlike the logistic map, this map is antisymmetric,
  that is, $f(-x) = -f(x)$.
%
In fact, all antisymmetric cubic maps \cite{brown3, brown4}
  can be mapped to \refeq{cubic} by a change of variables.
We shall show below that this odd parity can be used to reduce the amount of calculation.

@@

%Similar to the logistic-map case,
Even cube-free cyclic polynomials
can be mapped to \emph{even ternary} necklaces,
in which each bead can assume
0, 1, or 2, with the total being even.
%
The number of ternary necklaces is
$N_3(n) = \frac{1}{n} \sum_{d|n} \varphi(n/d) 3^d$,
%\begin{theorem}
while the even ones sum to
%or the cube-free cyclic polynomials for the cubic map of even degrees in $\vx$
%is given by
\begin{equation}
  N_e(n) = \frac{1}{n} \sum_{c d = n} \varphi(c)
    \left[
      3^d - \odd(c) \frac{3^d-1}{2}
    \right],
\label{eq:cubnecklace}
\end{equation}
where
$\odd(c)$ %$\equiv \big[1 - (-1)^c\big]/2$
is 1 if $c$ is odd or 0 if even.
%\label{thm:cubnecklace}
%\end{theorem}
%
Since $N_e \approx N_3/2$, using the parity saves about half of the equations
  in computation.
%
$N_e(n) = $ 2, 4, 6, 14, 26, 68, 158, 424, \dots, from $n = 1$.
%

@@

Finally, we mention that the $n$-cycle polynomial $P_n\rlam$
  is an even polynomial of $r$ if $n$ is a multiple of 4.
%This follows from that the transformation
%  $x_k \equiv (-)^k i y_k, r \equiv -r'$
%  yields the identical map: $y_{k+1} = r' y_k - {y_k}^3$.
Practically, we can cut the computating time by half
  when calculating the determinant by Lagrange interpolation
  in this case.

@@

@@

@@

>>>>

We will illustrate the algorithm on the logistic map% \cite{may, strogatz}
  defined in \refeq{logmap},
  although it can be readily generalized.
%
If $r$ is real,
  we will find windows $(r_a, r_b)$,
    within which stable $n$-cycles can exist.
%
%The window boundaries can be classified
%  to onset and bifurcation points according to \refeq{der};
Here, if $r > 0$, then
  $r_a$ ($r_b$) are the onset (bifurcation) points.
%
There are generally multiple such windows for a single $n$,
  but all onset points satisfy the same polynomial equation,
  and all bifurcation points another.
%
Our goal is thus to find the two polynomials for a given $n$.

@@

First, to count the square-free cyclic polynomials,
  we notice a one-to-one mapping
  from the square-free cyclic polynomials
  to the binary necklaces.
%
A binary necklace
  is a nonequivalent binary $0$-$1$ string.
%
Two strings are equivalent if they differ only by a circular shift.
%
For example, for $n = 3$, there are $2^3 = 8$ binary strings,
but only four necklaces: 000, 001, 011 and 111,
since 010 and 100 are equivalent to 001,
so are 110 and 101 to 011.
%
The mapping is constructed as follows.
%
Given a cyclic polynomial, if the generator contains $x_k$,
  the $k$th character from the right of the necklace is 1,
  otherwise 0,
e.g. the above four necklaces correspond to
  $C_0$, $C_1$, $C_{12}$ and $C_{123}$, respectively.
%For example, if $n = 3$,
% the square-free polynomials for $000$, $100$, $110$ and $111$ are
%  $C_0 = 1$ (generator: 1),
%  $C_1 = x_1 + x_2 + x_3$ (generator: $x_1$),
%  $C_{12} = x_1 x_2 + x_2 x_3 + x_3 x_1$ (generator: $x_1 x_2$)
%  and $C_{123} = x_1 x_2 x_3$ (generator: $x_1 x_2 x_3$), respectively.
Thus,
%
%\begin{theorem}
the number $\NB(n)$ of the square-free cyclic polynomials $C_p$,
  %$(\vx)$ formed by the cycle points $\vx = \{x_1, \ldots, x_n\}$,
  is given by \cite{riordan}
\begin{equation}
  %\deg_\lambda A_n\lam =
  \NB(n) = (1/n) \textstyle\sum_{d|n} \varphi(n/d) 2^{\,d},
\label{eq:necklace}
\end{equation}
%
%\label{thm:necklace}
%\end{theorem}
%
where $\varphi(m)$ is Euler's totient function,
which gives the number of integers from 1 to $m$
  that are coprime to $m$,
  e.g.,
  $\varphi(1) = 1$,
  $\varphi(3) = 2$ for 1 and 2,
  $\varphi(6) = 2$ for 1 and 5.
By \refeq{xr},
  $\NB$ is also $\deg_\lambda A_n$.
The first few $\NB(n)$ are
2, 3, 4, 6, 8, 14, 20, 36, 60, 108, \dots, % 188, 352, 632, \ldots,
  from $n = 1$.

@@

Incidentally, $\LB$ is also the number of aperiodic necklaces of length $n$
(e.g., for $n = 4$,
  ``$0011$'' is aperiodic, but ``$0101$'' is not due to the shorter period 2).
%
Since an $n$-necklace is either aperiodic itself or
  a repeated aperiodic $d$-necklace (with $d|n$),
  we have
\begin{equation}
  \NB(n) = \textstyle\sum_{d|n} \LB(d).
\label{eq:necklacelyndon}
\end{equation}
%
Since the left side is the degree in $\lambda$ of $A_n\lam$,
  which accommodates all fixed points of $f^n$,
  %\refeq{necklacelyndon} shows that
  each distinct $d$-cycle (with $d|n$)
  contributes one and only one zero to $A_n\lam$.
%
%
We may thus write
%\begin{theorem}
  \begin{equation}
    A_n\lam = \textstyle\prod_{d|n} Q_d^{n/d}\lam,
    \label{eq:prod}
  \end{equation}
where $Q_d^{n/d}(\lambda) \equiv \prod_l
  \big[\lambda - \Lambda\big(\vxn{n}^{(l)}\big) \big]$
  with $\vxn{n}^{(l)}$ being the $n$ points
  corresponding to the $l$th $d$-cycle.
%
In order words, $\vxn{n}^{(l)}$ is $\vxn{d}^{(l)}$
  repeated for $c = n/d$ times:
\begin{equation}
  \vxn{n}^{(l)} =
  \{
    x_1^{(l)}, \ldots, x_d^{(l)},
    \; \ldots, \;
    x_1^{(l)}, \ldots, x_d^{(l)}
  \}
  = \vxn{d}^{(l)} * c.
\notag
\end{equation}
%
It follows that
$\Lambda(\vxn{n})$ = $[\Lambda(\vxn{d})]^c$.
%
By factorizing $\lambda - \Lambda^c$
  in terms of the $c$ complex roots of $\lambda$,
  we get
%
\begin{align}
Q_d^{c}\lam
%  &= \prod_l
%    \Big\{
%      \lambda - \big[\Lambda\big(\vxn{d}^{(l)}\big) \big]^{c}
%    \Big\}
%    \notag \\
  &=
    \prod_l
    \Big\{
      \prod_{k=1}^c \,
      \big[
          e^{2k\pi i/c} \lambda^{1/c}
          - \Lambda\big(\vxn{d}^{(l)}\big)
      \big]
    \Big\},
    \notag
\end{align}
%
Now changing the order of the two products yields
\begin{equation}
Q_d^{c}\lam
  =  \prod_{k=1}^c P_d\big(e^{2k\pi i/c} \lambda^{1/c}\big),
\label{eq:QP}
\end{equation}
%
%
Despite the presence of $\lambda^{1/c}$,
  $Q_d^c\lam$ is a polynomial of $\lambda$,
  thanks to the invariance
  $\lambda \rightarrow e^{2\pi i} \lambda$;
and
  $\deg_\lambda Q_d^c\lam
    = \deg_\lambda P_d\lam = \LB(d)$.

@@

We can therefore distinguish \emph{original} cycles
  from those born out of the above branching.
Similar to \refeq{QP},
  the contributions from the non-original cycles
  born out of the $d$-cycles
  can be written as a cyclotomic product
  $\prod_{(k, c) = 1}
      P_{d}
        \left(
          R, e^{2k\pi i/c}
        \right)$,
where $k$ runs through integers from 1 to $c=n/d$,
  which are coprime to $c$.
%
This product for cycle intersection is equivalent to a previous
  formula in terms of canonical matrices \cite{blackhurst}.
%
With these contributions removed,
  we obtain the polynomial $S_n(R)$ responsible for the original cycles:
  %or the \emph{original factor} below,
  \begin{equation}
    S_n(R)
    = \frac
    {
      P_n(R, 1)
    }
    {
      \prod_{c d =  n, \; c > 1}
      %\left[ \,
        \prod_{(k, c) = 1}
      P_{d}
        \left(
          R, e^{2k\pi i/c}
        \right)
      %\right]
    },
  \label{eq:origfac}
  \end{equation}
%
%
%
Take the $n = 4$ case as an example, we have
$\prod_{k=1,3} P_1(R, e^{k\pi i/2})
%=(-4R-2i-1)(-4R+2i-1)
=16R^2+8R+5$
(for $c = 4$),
$P_2(R, -1) = 4R - 5$
(for $c = 2$),
and
$P_4(R, 1)
= (4R - 5) (16 R^2 + 8 R + 5)
  \bigl[
    (4R - 3)^3 - 108
  \bigr]$.
Dividing $P_4(R, 1)$ by the two factors yields
$S_4(R) = (4R-3)^3 - 108$.
%  whose only real root $R_a=(3+\sqrt[3]{108})/4$ corresponds to
%  the onset of the original cycle.
Note $R_a' = 5/4$ is not original,
  for it is born out of period-doubling the 2-cycle.
The two cycles are compared in \reffigs{cobweb}(d) and (e).

@@

A simple two-dimensional extension of the logistic map
  is the H\'enon map \cite{henon}:
\begin{equation}
  x_{k+1} = 1 + y_k - a \, {x_k}^2, \quad
  y_{k+1} = b \, x_k.
\label{eq:henon}
\end{equation}
%
By a change of variables \cite{huang}
  $x_k^\mathrm{new} \equiv a \, x_k^\mathrm{old},
   y_k^\mathrm{new} \equiv a \, y_k^\mathrm{old}$,
we have, in terms of the new variables,
\begin{equation}
  x_{k+1} = a + y_k - {x_k}^2, \quad
  y_{k+1} = b \, x_k.
%\tag{$\ref{eq:henon}'$}
\label{eq:henons}
\end{equation}
%
%Since neither $a$ nor $b$ is changed during the transformation,
\refeq{henon} and \refeq{henons} share the
same onset and bifurcation points
in terms of $a$ and $b$.
%
With $b \rightarrow 0$ and $a \rightarrow R$,
we recover the logistic map \refeq{logmaps}.

@@

The counterpart of \refeq{der}
  can be found from the Jacobian of $n$th iterate of \refeq{henons}
\[
  \vct J^n
  \equiv
  \left(
    \begin{array}{ccc}
      \partial x_{n+1}/\partial x_1 & \partial x_{n+1}/\partial y_1 \\
      \partial y_{n+1}/\partial x_1 & \partial y_{n+1}/\partial y_1 \\
    \end{array}
  \right),
\]
and the magnitude of both eigenvalues of the matrix must be less than 1.
By the chain rule, we have
  $\vct J^n (x_1) = \vct J^1(x_1) \cdots \vct J^1(x_n)$,
  where
$\vct J^1(x_k)
  =
  \left(
    \begin{array}{cc}
      -2 x_k & 1 \\
      b & 0
    \end{array}
  \right)$.
Thus, the eigenvalues $\lambda$ of $\vct J^n$ can be found from
%
\begin{equation}
\big|\, \lambda \, \vct I - \vct J^n(x_1) \,\big|
  = \lambda^2 - \Theta(\vx) \,\lambda +(-b)^n = 0,
%\tag{$\ref{eq:der}'$}
\label{eq:henonder}
\end{equation}
%
where $\Theta(\vx)$ and $(-b)^n$
  are the trace and determinant of $\vct J^n(x_1)$,
  respectively \cite{huang, hitzl}.
The onset and bifurcation points
  correspond to $\lambda = +1$ and $-1$, respectively

@@

As an example of a higher-order %one-dimensional
map, we consider the following cubic map \cite{strogatz}
%
\begin{equation}
  x_{k + 1} = f(x_k) = r \, x_k - {x_k}^3.
\label{eq:cubic}
\end{equation}
%
Unlike the logistic map, this map is antisymmetric,
  that is, $f(-x) = -f(x)$.
%
In fact, all antisymmetric cubic maps \cite{may2, brown3, brown4}
  can be mapped to \refeq{cubic} by a change of variables.
We shall show below that this odd parity can be used to reduce the amount of calculation.

@@

%Similar to the logistic-map case,
Even cube-free cyclic polynomials
can be mapped to \emph{even ternary} necklaces,
in which each bead can assume
0, 1, or 2, with the total being even.
%
The number of ternary necklaces is
$N_3(n) = \frac{1}{n} \sum_{d|n} \varphi(n/d) 3^d$,
%\begin{theorem}
while the even ones sum to
\begin{equation}
  N_e(n) = \frac{1}{n} \sum_{c d = n} \varphi(c)
    \left[
      3^d - \odd(c) \frac{3^d-1}{2}
    \right],
\label{eq:cubnecklace}
\end{equation}
where
$\odd(c)$ %$\equiv \big[1 - (-1)^c\big]/2$
is 1 if $c$ is odd or 0 if even.
%\label{thm:cubnecklace}
%\end{theorem}
%
Since $N_e \approx N_3/2$, using the parity saves about half of the equations
  in computation.
%
$N_e(n) =$ 2, 4, 6, 14, 26, 68, 158, 424, \dots, from $n = 1$.
%

@@

Finally, we mention that the $n$-cycle polynomial $P_n\rlam$,
  after divided by the contribution from the odd-cycles,
  is an even polynomial of $r$ for an even $n$.
%
This follows from that the transformed variables
  $y_k \equiv (-)^k i x_k$
satisfy
  $y_{k+1} = (-r) y_k - {y_k}^3$.
Thus, there is a one-to-one correspondence
  between the cycles for $r$ and $-r$.
%
Since the mapping applies to an even $m$-odd-cycle,
  hence the corresponding $2m$-cycle,
  $P_n\rlam$ is an even polynomial of $r$
  if $n$ is a multiple of 4.
%
This fact allows one to cut the computating time by half
  in calculating the determinant by Lagrange interpolation
  in these cases.

@@

\bibitem[Riordan(1958)]{riordan}
  Riordan, J. [1958]
  {\it An Introduction to Combinatorial Analysis}
  (Wiley, New York).

@@

\bibitem[Huang(1985)]{huang}
  Huang, Y.-N. [1985]
  ``Determination of the stable periodic orbits for the Henon map by analytical method,''
  {\it Chin. Phys. Lett.}
  \textbf{2},
  97--100.

@@

\bibitem[May(1979)]{may2}
  May, R. M. [1979]
  ``Bifurcations and dynamic complexity in ecological systems,''
  {\it Ann. N. Y. Acad. Sci.}
  \textbf{316},
  517--529.

### 2013-05-15 doc/pc.tex
add a few references, hengb.magma scripts
sightly improve doc

Back to our problem, there is
  a correspondence between the binary necklaces
  and the square-free cyclic polynomials.
For a square-free cyclic polynomial,
  we construct a binary string according to its generator:
if it contains $x_k$, the $k$th character from the left is 1,
  otherwise 0.
The resulting string corresponds to a unique necklace;
  the alternative generators give
  the circularly-shifted binary strings.
The mapping is reversible, or one-to-one.
For example, if $n = 3$,
 the square-free polynomials for $000$, $100$, $110$ and $111$ are
  $C_0 = 1$ (generator: 1),
  $C_1 = x_1 + x_2 + x_3$ (generator: $x_1$),
  $C_{12} = x_1 x_2 + x_2 x_3 + x_3 x_1$ (generator: $x_1 x_2$)
  and $C_{123} = x_1 x_2 x_3$ (generator: $x_1 x_2 x_3$), respectively.
\reftab{sqrfreepoly} shows a few more examples.
%
Thus,
%
\begin{theorem}
The number $\NB(n)$ of the square-free cyclic polynomials $C_p(\vx)$ ($p \in \Pset$)
  formed by $\vx = \{x_1, \ldots, x_n\}$ is
\begin{equation}
  \NB(n) = N(n) = (1/n) \textstyle\sum_{d|n} \varphi(n/d) 2^{\,d},
\label{eq:necklace}
\end{equation}
%
which is also equal to $\deg_\lambda A_n\Rlam$.
\label{thm:necklace}
\end{theorem}
%
Here, $\varphi(m)$ is Euler's totient function,
which gives the number of integers from 1 to $m$
  that are coprime to $m$,
  e.g.,
  $\varphi(1) = 1$,
  $\varphi(3) = 2$ for 1 and 2,
  $\varphi(6) = 2$ for 1 and 5.
The first few values of $N_\Pset(n)$ are 2, 3, 4, 6, 8, 14, 20, 36, 60, 108, 188, 352, 632, \ldots,
  from $n = 1$.

@@

A simple two-dimensional extension of the logistic map
  is the H\'enon map \cite{henon}:
\begin{equation}
  x_{k+1} = 1 + y_k - a \, {x_k}^2, \quad
  y_{k+1} = b \, x_k.
\label{eq:henon}
\end{equation}
%
By a change of variables
  $x_k^\mathrm{new} \equiv a \, x_k^\mathrm{old},
   y_k^\mathrm{new} \equiv a \, y_k^\mathrm{old}$,
we have, in terms of the new variables,
\begin{equation}
  x_{k+1} = a + y_k - {x_k}^2, \quad
  y_{k+1} = b \, x_k.
\tag{$\ref{eq:henon}'$}
\label{eq:henons}
\end{equation}
%
%Since neither $a$ nor $b$ is changed during the transformation,
\refeq{henon} and \refeq{henons} share the
same onset and bifurcation points
in terms of $a$ and $b$.
%
With $b \rightarrow 0$ and $a \rightarrow R$,
\refeq{henons} is reduced to the logistic map \refeq{logmaps}.

@@

We now study the following cubic map \cite{strogatz}
\begin{equation}
  x_{k + 1} = f(x_k) = r \, x_k - {x_k}^3.
\label{eq:cubic}
\end{equation}
This map is closely related to the one studied by Brown \cite{brown3}.
Since the new replacement rule
\begin{equation}
  {x_k}^3 \rightarrow r \, x_k - x_{k+1}
\label{eq:cubreplace}
\end{equation}
no longer eliminates squares,
%
we must extend the basis set of cyclic polynomials
  from the square-free ones to the cube-free ones,
  in using \refeq{xcp}.
We include in the basis of expansion
  $C_{112} = {x_1}^2 x_2 + {x_2}^2 x_3 + \dots + {x_n}^2 x_1$ ($n\ge3$),
but not
  $C_{1112} = {x_1}^3 x_2 + {x_2}^3 x_3 + \dots + {x_n}^3 x_1$.

@@

@@

@@

>>>>

Back to our problem, there is
  a correspondence between the binary necklaces
  and the square-free cyclic polynomials.
For a square-free cyclic polynomial,
  we construct a binary string according to its generator:
if it contains $x_k$, the $k$th character from the left is 1,
  otherwise 0.
The resulting string corresponds to a unique necklace;
  the alternative generators give
  the circularly-shifted binary strings.
The mapping is reversible, or one-to-one.
%
For example, if $n = 3$,
 the square-free polynomials for $000$, $100$, $110$ and $111$ are
  $C_0 = 1$ (generator: 1),
  $C_1 = x_1 + x_2 + x_3$ (generator: $x_1$),
  $C_{12} = x_1 x_2 + x_2 x_3 + x_3 x_1$ (generator: $x_1 x_2$)
  and $C_{123} = x_1 x_2 x_3$ (generator: $x_1 x_2 x_3$), respectively.
\reftab{sqrfreepoly} shows a few more examples.
%
Thus, we have \cite{riordan}
%
\begin{theorem}
The number $\NB(n)$ of the square-free cyclic polynomials $C_p(\vx)$ ($p \in \Pset$)
  formed by $\vx = \{x_1, \ldots, x_n\}$ is
\begin{equation}
  \NB(n) = N(n) = (1/n) \textstyle\sum_{d|n} \varphi(n/d) 2^{\,d},
\label{eq:necklace}
\end{equation}
%
which is also equal to $\deg_\lambda A_n\Rlam$.
\label{thm:necklace}
\end{theorem}
%
Here, $\varphi(m)$ is Euler's totient function,
which gives the number of integers from 1 to $m$
  that are coprime to $m$,
  e.g.,
  $\varphi(1) = 1$,
  $\varphi(3) = 2$ for 1 and 2,
  $\varphi(6) = 2$ for 1 and 5.
The first few values of $N_\Pset(n)$ are 2, 3, 4, 6, 8, 14, 20, 36, 60, 108, 188, 352, 632, \ldots,
  from $n = 1$.

@@

A simple two-dimensional extension of the logistic map
  is the H\'enon map \cite{henon}:
\begin{equation}
  x_{k+1} = 1 + y_k - a \, {x_k}^2, \quad
  y_{k+1} = b \, x_k.
\label{eq:henon}
\end{equation}
%
By a change of variables \cite{huang}
  $x_k^\mathrm{new} \equiv a \, x_k^\mathrm{old},
   y_k^\mathrm{new} \equiv a \, y_k^\mathrm{old}$,
we have, in terms of the new variables,
\begin{equation}
  x_{k+1} = a + y_k - {x_k}^2, \quad
  y_{k+1} = b \, x_k.
\tag{$\ref{eq:henon}'$}
\label{eq:henons}
\end{equation}
%
%Since neither $a$ nor $b$ is changed during the transformation,
\refeq{henon} and \refeq{henons} share the
same onset and bifurcation points
in terms of $a$ and $b$.
%
With $b \rightarrow 0$ and $a \rightarrow R$,
\refeq{henons} is reduced to the logistic map \refeq{logmaps}.

@@

We now study the following cubic map \cite{strogatz}
\begin{equation}
  x_{k + 1} = f(x_k) = r \, x_k - {x_k}^3.
\label{eq:cubic}
\end{equation}
This map is closely related to the one studied by Brown \cite{may2, brown3}.
Since the new replacement rule
\begin{equation}
  {x_k}^3 \rightarrow r \, x_k - x_{k+1}
\label{eq:cubreplace}
\end{equation}
no longer eliminates squares,
%
we must extend the basis set of cyclic polynomials
  from the square-free ones to the cube-free ones,
  in using \refeq{xcp}.
We include in the basis of expansion
  $C_{112} = {x_1}^2 x_2 + {x_2}^2 x_3 + \dots + {x_n}^2 x_1$ ($n\ge3$),
but not
  $C_{1112} = {x_1}^3 x_2 + {x_2}^3 x_3 + \dots + {x_n}^3 x_1$.

@@

\bibitem{riordan}
  J. Riordan
  \textit{An Introduction to Combinatorial Analysis}
  (Wiley, New York, 1958).

@@

\bibitem{huang}
  Y. N. Huang
  Determination of the stable periodic orbits for the Henon map by analytical method,
  \textit{Chin. Phys. Lett.}
  \textbf{2},
  97--100
  (1985).

@@

\bibitem{may2}
  R. M. May,
  Bifurcations and dynamic complexity in ecological systems,
  \textit{Ann. N. Y. Acad. Sci.}
  \textbf{316},
  517--529
  (1979).
